---
title: "Assignment_1STA"
author: "Calvin Gacheru"
date: "`r Sys.Date()`"
output: html_document
---

## Generating the EUStockMarkets data set:

*loading the data set to R:*

```{r echo = FALSE}
datasets::EuStockMarkets
View(EuStockMarkets)  # for viewing the data set in a different tab.

```

### Main focus of EuStockMarkets data set:

The European Stock Markets data set focuses on the daily closing of major European stock indices: Germany DAX (Ibis), Switzerland SMI, France CAC, and UK FTSE. The data set shows a financial market theme for data collected on a daily basis from 1991-1998 excluding weekends and holidays.

### Intended use of the EuStockMarkets data set:

The EuStockMarkets is used to provide statistical analysis in the European Stock Market, compare different stocks from the four European Countries and develop predictive analysis for the stocks.

### Structure and organization of EuStockMarkets data set:

The data set provides a multivariate time series with 1860 observation on 4 variables. The four variables represent: Germany DAX (Ibis) Switzerland SMI France CAC UK FTSE.

*generating a summary for all numerical variables: using the summary function.*

```{r}
summary(EuStockMarkets) 
```

*The summary function provides a minimum, 1st and 3rd quartile, mean, median and maximum of DAX, SMI, CAC and FTSE. To summarize individual variables i.e Germany DAX:*

```{r}
summary(EuStockMarkets[, c('DAX')])
summary(EuStockMarkets[, c('SMI')])
summary(EuStockMarkets[, c('CAC')])
summary(EuStockMarkets[, c('FTSE')])

```

*using describe function:*

```{r}
library(psych)   #loading the psych package to enable describe function
describe(EuStockMarkets)
```

*Calculating mean of the four stocks:*

```{r}
mean(EuStockMarkets[,'DAX'])
mean(EuStockMarkets[,'SMI'])
mean(EuStockMarkets[,'CAC'])
mean(EuStockMarkets[,'FTSE'])
```

The mean are as follows DAX = 2530.657

SMI = 3376.224

CAC = 2227.828

FTSE = 3565.643

The mean provides an average daily closing from 1991-1998. It can also used in predictive modeling and other statistical analysis.

*Calculating median of the four stocks:*

```{r}
median(EuStockMarkets[,'DAX'])
median(EuStockMarkets[,'SMI'])
median(EuStockMarkets[,'CAC'])
median(EuStockMarkets[,'FTSE'])

```

The median are as follows: DAX = 2140.565

SMI = 2796.35

CAC = 1992.3

FTSE = 3246.6

The median on the other hand represents the middle daily closing index from 1991-1998. Its significance is noted in its robustness since it is not affected by extreme values hence can provide a more accurate value in the data set.

*visualizing the EuStockMarkets with a line plot:*

```{r}
library(ggfortify)
library(ggplot2)
autoplot(EuStockMarkets) #automatically plot the EuStockMarkets
```

### Noticeable patterns and trends for the stock index

From the visualization above, we can take note the following trends:

1)  When one index increases, the other indexes tend to also increase, for example, when the stock index of Germany DAX increases, the three other stocks indexes also increase.

2)  The stock index of all four DAX, SMI, CAC and FTSE fluctuates from 1991-1998. It does not rise nor fall steadily, rather it rises then falls, then rises.

### Distribution of the data and insights that can be drawn from it.

The distribution of the EustockMarkets can provide insights such as measuring the mean and the median to get the average daily closing of each stock index since 1991-1998, excluding weekends and holidays.
